---
title: "Chapter4_WorkflowBasics"
author: "Ryann Crowley"
date: "March 24, 2017"
output: word_document
output: md_document
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

#4.1 Coding basics

You can use R as a calculator.  
You can create objects with <-.
Keyboard shortcut for <- is Alt+- (the minus sign). Oh thank goodness!! Must commit to muscle memory. :) 


#4.2 what's in a name

object names must start with a letter and can only contain letters, numbers and .

Rstudio has an auto-completion facility.  Start by typing an object name, add characters until you have a unique prefix, then press return or tab.  Example below.
```{r}
this_is_a_really_long_name <- 2.5
```

```{r}
this_is_a_really_long_name
```

Oops, the variable above should have a value of 3.5.  Stary by typing "this" and reassign using the auto-completion and the Ctrl+ Up arrow shortcut. The new shortcut will list all the commands you've typed taht start with those letters.

```{r}
this_is_a_really_long_name <- 3.5
```