---
title: "Chapter 2 - Visualization"
author: "Ryann Crowley"
date: "March 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
.libPaths()
```

```{r}
install.packages(scales)
```

```{r}
.packages(all = TRUE)
```

```{r}
library(scales)
library(tidyverse)
```

```{r}
mpg
```


Do larger cars use more gas?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

The plot shows a negative relationship between engine size (displ) and fuel efficiency (hwy)


Turns plot code intoa reusable template for making graphs wiht ggplot2. 

```{r}
ggplot(data = mpg)
```



```{r}
str(mtcars)
head(mpg)
```
```{r}
?mpg
```

Scatterplot of class ves drv

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```
this plot is not helpful at all!

Going back to hwy by displ (mileage by engine type), let's also look at class of car (class) by adding it to the aes (asthetic statement)

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

The plot reveals that many of the outliers are two seater cars.  In the plot above, we mapped class to teh szie aesthetic. We could also assing alpha (shading), shape, or size.

We can also assign our own aesthetic.  For example, set the points to blue. Notice, the "color = blue" statemetn is AFTER the paratheses of the aes statement.

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

##Exercises 3.3
2. Which variables in mpg are categorical? Which are continuous?

```{r}
?mpg
str(mpg)
```

3. Map a continous variable to color, size, and shape.  How do these aesthetics behave differently for categorical vs. continous variables?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = year, size = cty))
```
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = year, shape = cty))
```
Looks like continous variables cannot be assigned to shape.

4. What happens when you map the same variable to multiple aethetics?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = year, size = year))
```
Looks like the last aesthetic designation is the only one that is accurately displayed.  The color aesthetic has a legend but it not accurately portraded.

5. What does the stroke aesthetic do?  What shapes does it work with?

```{r}
??stroke
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, stroke = year))
```
I don't really have a good answer for this.

6. What happens if you map an aesthetic to something other than a variable name, like aes(color = displ<5)

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = displ<5))
```
Very cool.  Specific select statements can be made within the aesthetic!

#Section 3.4 Common Problems
#Section 3.5 Facets

Facets a plots split by categorical variables using "facet_wrap()".

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```

To facet your plots with a combination of two variables, use facet_grid()

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(drv ~ cyl)
```
If you prefer not to facet a row or a column us the . character.
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(. ~ cyl)
```

##Exercises 3.5
1. What happens if you facet a continous variable?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ cty)
```
Makes a graph for EVERY value.

2. What do the empty cells in plot wiht facet_grid(drv ~ cyl) mean? How do they relate to this plot?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = cyl)) 
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(drv ~ cyl)
```
Well, if cell value in the crosstabs of facet_grid is empty, there would be no data to display. Looks like the graph is omitted.  For exmample, the cross of drv = r and cyl = 4. 

3. What plots does the following code make?  What does the . do?

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(. ~ cyl)
```
The . designates "ignore". If you specify grid but don't only have one varialbe, you can fill the other with a period.  One advantage, you can specify if you want the faceting to create multiple rows or columns. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_grid(cyl ~ .)
```
4. Take the first faceted plot in this section: (see below).  What are the advanges of using faceting instead of color aesthetic?  What are the disadvantages?  How might the balance change if you had a larger dataset?


```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + facet_wrap(~ class, nrow = 2)
```
Advantage - easier to see differences within a specific factet.  Disadvantage - harder to see overall trends across all data.  With a larger dataset, facetes might be even more useful to sift trhough the large number of records.

5. Read ?facet_wrap.  What does nrow do? What does ncol do?  What other options control the layout of the individual panels?  Why doesn't facet_grid() have nrow and ncol?

```{r}
?facet_wrap
```

nrow and ncol - disignate the number of rows and columns. facet_grid - let's the data designate the number of rows and columns. as.table - if TRUE, will lay out the facets like a table with higest avleus at teh bottom-right.  If FALSE, will lay out like a plot with highest values at the top-right. drop, if TRUE, all factors level not used in teh data will automatically be dropped.

6. When using facet_grid() you should usually put the variable with more unique levels in the columns.  Why?

Easier to compare side by side versus top to bottom?

#Section 3.6 Geometric objects

Geoms are the geometric objects used to produce the plot.  For example: Blox, line, smooth, point. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

Not every aesthetic works with every geom.  For example, linetype does not work for geom_point...but linetype does work with smooth.

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))
```
To make the linetype stand out even more, it can also use color.  You can even overlay TWO geoms in the same graph. We will learn how to do so shortly.

There are over 30 geoms in ggplot2.  Get a cheatsheet!

For geoms that create a single object for multiple records (rows of data), a seperate object can be create for each GROUP.

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

To display multiple geoms in the same plot, add mutiple geom functions to ggplot().

```{r}
ggplot(data = mpg) + 
  geom_smooth(mapping = aes(x = displ, y = hwy)) + geom_point(mapping = aes(x = displ, y = hwy))
```

The following will produce teh same graph while only specifying the data once.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```

You can add different aesthetics to different layers.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```
Add you can add different data to different layers. In this example, the smooth geom is only for subcompact cars.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)
```

##Exercises 3.6
1. What geom would you use to draw a line chart? boxplot? histogram? area chart?
geom_line, geom_boxplot, geom_histogram, geom_area

2. Run teh followign, what does se = FALSE do?
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Removes the Standard Error around the lines.

3. What does show.legend = FALSE do?  What happens if you remove it?

Remove legend.  It is the default. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, line = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, line = drv, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
``


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv) + 
  geom_smooth(mapping = aes(x = displ, y = hwy)
```
```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy), se = FALSE)
```
```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv), se = FALSE)
```
